### 2016-02-01 papers/CAFe/HIPS-paper/intro.tex
tweaks

Another potential option is to abandon the stricture of message passing within a procedural
programming language altogether.  For example, the Open Community Runtime (OCR)
project\cite{OCR:wiki:url} provides an asynchronous task-based runtime designed for
machines with high-core count\cite{Dokulil20151453}, such as can be expected in any
exascale hardware design.  The project's goal is to create a framework and reference
implementation to help developers explore programming methods to improve the power
efficiency, programmability, and reliability of HPC applications while maintaining
application performance.

@@

It cannot be emphasized too strongly that using the OpenCL programming language and
libraries (or programming for the OCR environment) presents a high hurdle for the
application developer creating real scientific applications.  A goal of this paper is to
examine ways in which high-level syntax based on Fortran coarrays can be utilized by
compiler-based tools to \emph{target} programming environments like OpenCL, thus
freeing the developer from the task of frequently porting the application to new complex
environments.  The hope is that high-level syntax can hide the complexities of the
underlying hardware \emph{and software.}

@@

Since each CAF process executes the same program, there are no good avenues for exploiting
task-based parallelism.  Neither is there a adequate mechanism for dynamic coarray memory
allocation, nor a mechanism for the remote execution of functions.  In CAF, like MPI, the
programmer often combines distributed memory processes running the same program image with a
shared memory threaded programming model using OpenMP (or OpenACC).  This model breaks down
when dealing with architectures that are not isomorphic to the model for which MPI and
relatives were designed -- a distributed collection of homogenous processing elements in
which each processing element supports a small number of parallel threads of execution.  For
example, an HPC system that makes use of traditional multicore CPUs as well as GPGPU accelerators
breaks this model and, to effectively program it, currently requires a mixture of programming
systems in a single application.

>>>>

Another potential option is to abandon the structure of message passing within a procedural
programming language altogether.  For example, the Open Community Runtime (OCR)
project\cite{OCR:wiki:url} provides an asynchronous task-based runtime designed for
machines with high-core count\cite{Dokulil20151453}, such as can be expected in any
exascale hardware design.  The project's goal is to create a framework and reference
implementation to help developers explore programming methods to improve the power
efficiency, programmability, and reliability of HPC applications while maintaining
application performance.

@@

It cannot be emphasized too strongly that using the OpenCL programming language and
libraries (or programming for the OCR environment) presents a high hurdle for the
application developer creating real scientific applications.  A goal of this paper is to
examine ways in which high-level syntax based on Fortran coarrays can be utilized by
compiler-based tools to \emph{target} programming environments like OpenCL, thus
freeing the developer from the task of frequently porting the application to new complex
environments.  The hope is that high-level syntax can hide the complexities of the
underlying hardware \emph{and software} by delegating some optimization tasks to the
compiler instead of the programmer.

%% MJS: this seems like a natural place for a subsection break, but it looks ugly
%%      in the IEEE two column style

@@

Since each CAF process executes the same program, there are no good avenues for exploiting
task-based parallelism.  Neither is there an adequate mechanism for dynamic coarray memory
allocation, nor a mechanism for the remote execution of functions.  In CAF, like MPI, the
programmer often combines distributed memory processes running the same program image with a
shared memory threaded programming model using OpenMP (or OpenACC).  This model breaks down
when dealing with architectures that are not isomorphic to the model for which MPI and
relatives were designed -- a distributed collection of homogenous processing elements in
which each processing element supports a small number of parallel threads of execution.  For
example, an HPC system that makes use of traditional multicore CPUs as well as GPGPU accelerators
breaks this model and, to effectively program it, currently requires a mixture of programming
systems in a single application.

### 2016-02-01 papers/CAFe/HIPS-paper/performance.tex

\begin{comment}
However, the relative performance of computation on a cluster of GPUs compared with the
necessary communication of halo information is of interest, especially considering that a
complete exchance of halo data involves communication between multiple coarray images
\emph{and} between each individual host image and its subimage (an attached OpenCL
device).
\end{comment}

>>>>

\begin{comment}
However, the relative performance of computation on a cluster of GPUs compared with the
necessary communication of halo information is of interest, especially considering that a
complete exchange of halo data involves communication between multiple coarray images
\emph{and} between each individual host image and its subimage (an attached OpenCL
device).
\end{comment}

### 2016-02-01 papers/CAFe/HIPS-paper/transformations.tex

These transformations are accomplished using several rewrite stages using Stratego/XT
strategies:
(1) a symbol table is produced in the first pass to store information related to arrays
including, array shape and allocation status;
(2) additional variables are declared that are used to maintain the OpenCL runtime state,
including the OpenCL device, the OpenCL kernel, and OpenCL variables used to marshall
kernel arguments; and
(3) all CAFe code related to subimage usage is desugared (lowered) to standard Fortran
with calls to the ForOpenCL library as needed.
The latter step includes the allocation of OpenCL device memory, transfer of memory to and
from the OpenCL device, marshalling of kernel arguments, running of the OpenCL kernel, and
synchronization.

>>>>

These transformations are accomplished using several rewrite stages using Stratego/XT
strategies:
(1) a symbol table is produced in the first pass to store information related to arrays,
including array shape and allocation status;
(2) additional variables are declared that are used to maintain the OpenCL runtime state,
including the OpenCL device, the OpenCL kernel, and OpenCL variables used to marshall
kernel arguments; and
(3) all CAFe code related to subimage usage is desugared (lowered) to standard Fortran
with calls to the ForOpenCL library as needed.
The latter step includes the allocation of OpenCL device memory, transfer of memory to and
from the OpenCL device, marshalling of kernel arguments, running of the OpenCL kernel, and
synchronization.
